---
title: "R for Data Science Notes"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```

# Chapter 3 - Data visualisation
Stat summary looks useful

```{r }
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```

And geom_smooth

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth()
```

# Chapter 5 - Data transformation
Use is.na() for NA - I sometimes forget to do this
```{r}
df <- tibble(x = c(1, NA, 3))
filter(df, x > 1)
filter(df, is.na(x) | x > 1)
```

starts_with("abc"): matches names that begin with “abc”.

ends_with("xyz"): matches names that end with “xyz”.

contains("ijk"): matches names that contain “ijk”.

matches("(.)\\1"): selects variables that match a regular expression. This one matches any variables that contain repeated characters. You’ll learn more about regular expressions in strings.

num_range("x", 1:3): matches x1, x2 and x3
rename(flights, tail_num = tailnum)

Use arrange to order rows - I will use this in future rather than base version

```{r flights}
arrange(flights, year, month, day)
```

Use desc() to re-order by a column in descending order
NA are at the end

```{r}
df <- tibble(x = c(5, 2, NA))
arrange(df, x)
arrange(df, desc(x))
```

group_by works nicely with summarise

```{r}
delays <- flights %>% 
  group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
  filter(count > 20, dest != "HNL")
```

Fortunately, all aggregation functions have an na.rm argument which removes the missing values prior to computation
NAs are contagious

```{r}
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay, na.rm = TRUE))
```


